A Data-Driven Framework for Neural Field Modelling

### 2010-11-02 ltx/first_order_framework/LaTeX_Manuscript.tex
minor correction

\subsubsection{Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where \dean{150} realisations of the data were generated. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure the parameters  converged. Figure~\ref{fig:Figure6} shows the rate of convergence, confirming the parameters converged, with the changes in the parameter errors falling to less than $10^{-4}$ after six iterations.

Histograms of the parameter estimates obtained over the \dean{150} realisations are shown in Figure~\ref{fig:Figure7}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good agreement with the actual values of the connectivity kernel parameters, however the time constant parameter, $\xi$, is outside the estimate distribution (with bias of 2.78\%).

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for \dean{150} realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around 0.5~mV, which is approximately 7.1~\% of the field voltage range ([-3.29,3.77]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(b) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(a)), but is missing the higher-frequency detail.

@@

To further study the performance of the proposed framework we used different sets of connectivity basis functions parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Figure~\ref{fig:Figure8}. For each subset of parameters the mean of the reconstructed kernel over \dean{50} realisations is shown as a dotted line and the 95\% confidence interval is indicated by the grey shaded region.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure6.pdf}
\end{center}
\caption{{\bf Convergence of parameters.} The mean of the absolute error between true and estimated parameters averaged over \dean{150} realisations. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:Figure6}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure7.pdf}
\end{center}
\caption{{\bf Histograms of the parameter estimates over \dean{150}
realisations}. The solid lines show the actual values of the parameters and the dotted lines show the means of the estimated parameters. (\textbf{a}) The histogram of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. (\textbf{b}) The histogram of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. (\textbf{c}) The histogram of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. (\textbf{d}) The histogram of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Figure7}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure9.pdf} 
\end{center}
\caption{{\bf Error in the field reconstruction}. The mean of average RMSE of the estimated field over \dean{150} realisations (solid line) and the 95~\% confidence interval (grey region).} 
\label{fig:Figure9}
\end{figure}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure8_1.pdf}
\end{center}
\caption{{\bf Cross-section of connectivity kernel estimate}. The actual kernel is shown by the solid line, the mean kernel estimate (over \dean{50} realisations) is shown by the dotted line and the 95~\% confidence interval is shown by the shaded grey region.}
\label{fig:Figure8}
\end{figure}

>>>>

\subsubsection{Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realisations of the data were generated. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure the parameters  converged. Figure~\ref{fig:Figure6} shows the rate of convergence, confirming the parameters converged, with the changes in the parameter errors falling to less than $10^{-4}$ after six iterations.

Histograms of the parameter estimates obtained over the 150 realisations are shown in Figure~\ref{fig:Figure7}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good agreement with the actual values of the connectivity kernel parameters, however the time constant parameter, $\xi$, is outside the estimate distribution (with bias of 2.78\%).

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around 0.5~mV, which is approximately 7.1~\% of the field voltage range ([-3.29,3.77]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(b) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(a)), but is missing the higher-frequency detail.

@@

To further study the performance of the proposed framework we used different sets of connectivity basis functions parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Figure~\ref{fig:Figure8}. For each subset of parameters the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95\% confidence interval is indicated by the grey shaded region.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure6.pdf}
\end{center}
\caption{{\bf Convergence of parameters.} The mean of the absolute error between true and estimated parameters averaged over 150 realisations. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:Figure6}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure7.pdf}
\end{center}
\caption{{\bf Histograms of the parameter estimates over 150 realisations}. The solid lines show the actual values of the parameters and the dotted lines show the means of the estimated parameters. (\textbf{a}) The histogram of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. (\textbf{b}) The histogram of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. (\textbf{c}) The histogram of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. (\textbf{d}) The histogram of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Figure7}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure9.pdf} 
\end{center}
\caption{{\bf Error in the field reconstruction}. The mean of average RMSE of the estimated field over 150 realisations (solid line) and the 95~\% confidence interval (grey region).} 
\label{fig:Figure9}
\end{figure}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure8_1.pdf}
\end{center}
\caption{{\bf Cross-section of connectivity kernel estimate}. The actual kernel is shown by the solid line, the mean kernel estimate (over 50 realisations) is shown by the dotted line and the 95~\% confidence interval is shown by the shaded grey region.}
\label{fig:Figure8}
\end{figure}
